---
title: "Results"
output: word_document
---

```{r libraries, include=FALSE}
library(here)
library(tidyverse)
library(testthat)
library(rstan)
library(psych)
library(GPArotation)
library(pander)
library(brms)
library(broom)
library(reshape2)
library(ggridges)
library(patchwork)
library(scales)
library(devtools)
library(tidybayes)
library(flextable)
library(officer)
```

```{r options, include=FALSE}
################################################################################
overwrite <- FALSE
overwrite_brms <- FALSE
################################################################################

n2_lower_bound <- 225
n2_upper_bound <- 325

p3_lower_bound <- 325
p3_upper_bound <- 625

seed <- 2020
cores <- 8
chains <- 8
iter <- 4000
warmup <- 1000
refresh <- 0

control_vs_patient <-
  scale_fill_manual(values = c("#386cb0", "#d95f02"),
                    labels = c("Controls", "Patients"))

erp_colors <-
  scale_color_manual(values = c("#386cb0", "#d95f02"),
                     labels = c("Controls", "MDOs"))

categorical <- c("gray", "black")
difference_color <- "#55C667"

axis_text_size <- 9
title_text_size <- 11
axis_title_size <- 12
subtitle_text_size <- 10
legend_text_size <- 10

dir.create(here("/analysis/output"))
dir.create(here("/analysis/output/tables"))
dir.create(here("/analysis/output/figures"))
dir.create(here("/analysis/data/tmp"))
```

```{r helper functions, include=FALSE}
source(here("analysis/scripts/r/00_helper_functions.R"))
```

```{r brms models, include=FALSE}
source(here("analysis/scripts/r/01_brms_models.R"))
```

```{r read data, include=FALSE}
source(here("analysis/scripts/r/02_read_data.R"))
merged_data <- readRDS(here("analysis/data/tmp/merged_data.rds"))
```

## Materials and Methods

### Measures

#### Trait disinhibition 

```{r trait disinhibition reliability, include=FALSE}
if (!file.exists(here("analysis/data/tmp/gen_dis_reliability.rds")) || overwrite == TRUE) {
  merged_data <- readRDS(here("analysis/data/tmp/merged_data.rds"))
  gen_dis_reliability <-
    merged_data %>% 
    select(esi1, esi9, esi10, esi19, esi28,
           esi36, esi41, esi44, esi49, esi65,
           esi73, esi84, esi90, esi92, esi95,
           esi112, esi125, esi143, esi144, esi152) %>%
    omega(m = .,
          fm = "ml",
          n.factors = 1,
          plot = FALSE,
          lavaan = TRUE)
  
  gen_dis_reliability <- as.data.frame(gen_dis_reliability[c(3, 4)]) %>%
    mutate_if(is.numeric, round, 2)
  
  saveRDS(gen_dis_reliability, here("analysis/data/tmp/gen_dis_reliability.rds"))
}
  
gen_dis_reliability <- readRDS(here("analysis/data/tmp/gen_dis_reliability.rds"))
```

The ESI-BF~DIS~ showed high internal reliability in the current study, with Cronbach’s alpha = `r gen_dis_reliability$alpha` and McDonald’s Omega total = `r gen_dis_reliability$omega.tot`.

### Procedure

#### EEG data preprocessing

```{r group difference in preprocessing, include=FALSE}
source(here("analysis/scripts/r/03_robust_group_differences_preprocessing.R"))
```

`r length(remove)` participant in the patient group had less than four correct NoGo trials left after artifact rejection, and was excluded.

#### Reliability analysis

```{r reliability analysis, include=FALSE}
source(here("analysis/scripts/r/04_reliability_analysis.R"))
```

The reliability of the Go/NoGo paradigm was high, with ⍴~SB~ = `r reliability_models[2, 2]` for median NoGo response time and ⍴~SB~ = `r reliability_models[1, 2]` for NoGo accuracy. With the exception of NoGo N2 latency (⍴~SB~ = `r reliability_models[5, 2]`), the reliability of all ERP measurements was high, with ⍴~SB~ = `r reliability_models[3, 2]` for NoGo N2~WIN~, ⍴~SB~ = `r reliability_models[4, 2]` for NoGo N2~MOV~, ⍴~SB~ = `r reliability_models[6, 2]` for NoGo P3~WIN~, ⍴~SB~ = `r reliability_models[7, 2]` for NoGo P3~MOV~, and ⍴~SB~ = `r reliability_models[8, 2]` for NoGo P3 latency.

## Results

### Sample characteristics

#### Age and education

```{r age and education, echo=FALSE, message=FALSE, warning=FALSE}
if (!file.exists(here("analysis/data/tmp/group_differences_age.rds")) || overwrite == TRUE) {
  brms_robust_group_diff <- readRDS(here("analysis/scripts/brms/brms_robust_group_diff.rds"))
  group_diff_age <- get_group_difference(merged_data, "age")
  saveRDS(group_diff_age, here("analysis/data/tmp/group_differences_age.rds"))
}

age_summary <-
  readRDS(here("analysis/data/tmp/group_differences_age.rds")) %>%
  mutate(
      mean_whole_group = mean,
      range_whole_group = range,
      mean_controls = mean.1,
      range_controls = range.1,
      mean_patients = mean.2,
      range_patients = range.2) %>%
    select(-c(mean, range, mean.1, range.1, mean.2, range.2)) %>%
    mutate_if(is.numeric, round, 2) %>%
    mutate(diff = paste0(diff, " [", diff_lower_90, ", ", diff_upper_90, "]"),
           delta_est = paste0(delta_est, " [", delta_est_lower_90, ", ", delta_est_upper_90, "]"),
           Measure = variable) %>%
    select(Measure,
           mean_whole_group, range_whole_group,
           mean_controls, range_controls,
           mean_patients, range_patients,
           diff, delta_est, p_direction)

education_summary <- merged_data %>%
  group_by(group, education) %>%
  summarise(n = n()) %>%
  mutate(percent = round(n / sum(n) * 100, 0))

order <- education_summary %>%
  arrange(desc(n))

education_summary <- education_summary %>%
  mutate(percent = round(n / sum(n) * 100, 0),
         education = factor(education, unique(order$education)))

education_percent <- education_summary %>% 
             group_by(group) %>%
             filter(percent == max(percent)) %>%
             arrange(group)
``` 

Participants were on average `r strsplit(age_summary$mean_whole_group," ")[[1]][1]` years old (SD = `r strsplit(age_summary$mean_whole_group," ")[[1]][3]`, range = `r age_summary$range_whole_group`). MDOs were approximately `r round(as.numeric(strsplit(age_summary$diff," ")[[1]][1]), 0)` years older than controls, with an average of `r strsplit(age_summary$mean_patients," ")[[1]][1]` years (SD = `r strsplit(age_summary$mean_patients," ")[[1]][3]`, range = `r age_summary$range_patients`) for MDOs and `r strsplit(age_summary$mean_controls," ")[[1]][1]` years (SD = `r strsplit(age_summary$mean_controls," ")[[1]][3]`, range = `r age_summary$range_controls`) for controls. The estimated difference was `r age_summary$diff`, with $\hat{\delta}$ = `r age_summary$delta_est` and P~D~ = `r age_summary$p_direction`%. The most frequent level of education in both groups was having finished high school (`r as.numeric(education_percent[1, 4])`% in the control group and `r as.numeric(education_percent[2, 4])`% in the patient group), although `r as.numeric(education_summary[4, 3])` participants in the patient group (`r as.numeric(education_summary[4, 4])`%) had not finished primary school.

#### MDO group clinical characteristics

```{r clinical characteristics, message=FALSE, warning=FALSE, include=FALSE}
primary_dx <-
  merged_data %>%
  filter(group == "patient") %>%
  group_by(primary_diagnosis) %>%
  summarise(n = n()) %>%
  mutate(percent = round(n / sum(n) * 100, 0)) %>%
  arrange(desc(n))

additional_dx_summary <- 
  merged_data %>%
  filter(group == "patient") %>%
  summarize(n = n(),
            mean = round(mean(num_additional_dx), 2),
            median = median(num_additional_dx),
            sd = round(sd(num_additional_dx), 2),
            min = min(num_additional_dx),
            max = max(num_additional_dx))

additional_dx_list <-
  rbind(merged_data[, "additional_diagnosis_1"],
        merged_data[, "additional_diagnosis_2"],
        merged_data[, "additional_diagnosis_3"],
        merged_data[, "additional_diagnosis_4"])

additional_dx_list <- 
  sort(table(additional_dx_list), decreasing = TRUE) %>%
  as.data.frame() %>%
  mutate(percent = Freq / 27 * 100) %>%
  mutate_if(is.numeric, round, 0)

pharma_data <- subset(merged_data, group == "patient") %>%
  select(antipsychotics,
         antidepressants,
         sedatives_hypnotics_bz_bz_like,
         sedatives_hypnotics_non_bz_non_bz_like,
         anticholinergics,
         antiepileptics,
         adhd_substances,
         sud_substances,
         somatic_substances) %>%
  mutate(total_num_pharma = rowSums(., na.rm = TRUE))

pharma_data_sum <- 
  pharma_data %>%
    summarize(n = n(),
              mean = round(mean(total_num_pharma), 2),
              median = median(total_num_pharma),
              sd = round(sd(total_num_pharma), 2),
              min = min(total_num_pharma),
              max = max(total_num_pharma))
```

Overall, MDOs (N = `r c(table(merged_data$group))[[2]]`) presented with a wide range of mental disorders. The most common primary diagnosis was `r str_to_lower(primary_dx[1, 1])` (N = `r as.numeric(primary_dx[1, 2])`, `r as.numeric(primary_dx[1, 3])`%), while the most common additional diagnoses were `r str_to_lower(additional_dx_list[1, 1])` (N = `r additional_dx_list[1, 2]`, `r additional_dx_list[1, 3]`%) and `r str_to_lower(additional_dx_list[2, 1])` (N = `r additional_dx_list[2, 2]`, `r additional_dx_list[2, 3]`%). Most MDOs had `r additional_dx_summary$median` additional diagnosis (i.e., median = `r additional_dx_summary$median`), ranging from  `r additional_dx_summary$min` to `r additional_dx_summary$max`. See Figure X for a diagnostic overview. Most MDOs also received some form of pharmacological treatment (median = `r pharma_data_sum$median` different substances, range = `r pharma_data_sum$min` to `r pharma_data_sum$max`), although `r pharma_data %>% filter(total_num_pharma == 0) %>% nrow()` MDOs did not receive any pharmacological treatment at all. An overview of pharmacological treatment is presented in Table 1.

```{r pharmacological data, echo=FALSE, message=FALSE, warning=FALSE}
ft <-
  melt(pharma_data, measure.vars = c("antipsychotics",
                                   "antidepressants",
                                   "sedatives_hypnotics_bz_bz_like",
                                   "sedatives_hypnotics_non_bz_non_bz_like",
                                   "anticholinergics",
                                   "antiepileptics",
                                   "adhd_substances",
                                   "sud_substances",
                                   "somatic_substances")) %>%
  mutate(variable = gsub("antipsychotics", "Antipsychotics", variable),
         variable = gsub("antidepressants", "Antidepressants", variable),
         variable = gsub("sedatives_hypnotics_bz_bz_like", "Benzodiazepine sedatives/hypnotics", variable),
         variable = gsub("sedatives_hypnotics_non_bz_non_bz_like", "Non-benzodiazepine sedatives/hypnotics", variable),
         variable = gsub("anticholinergics", "Anticholinergics", variable),
         variable = gsub("antiepileptics", "Antiepileptics", variable),
         variable = gsub("adhd_substances", "ADHD substances", variable),
         variable = gsub("sud_substances", "SUD substances", variable),
         variable = gsub("somatic_substances", "Somatic substances", variable)) %>%
  group_by(variable, value) %>%
  summarise(n = n()) %>%
  mutate(percent = round((n / sum(n) * 100), 0)) %>%
  na.omit() %>%
  mutate(n = paste0(n, " (", percent, "%)")) %>%
  select(variable, n) %>%
  data.frame() %>%
  flextable() %>%
  set_header_labels(variable = "Pharmacological category",
                    n = "N (%)") %>%
  set_caption("Table 1. Overview of pharmacological treatment in the mentally disordered offenders group (N = 27).") %>%
  align(j = 1, align = "left", part = "all") %>%
  align(j = 2, align = "middle", part = "all") %>%
  autofit()

if (!file.exists(here("analysis/output/tables/table_1_pharmacological_treatment.docx")) || overwrite == TRUE) {
  read_docx() %>%
    body_add_flextable(value = ft, split = FALSE) %>%
    body_end_section_landscape() %>% # a landscape section is ending here
    print(target = here("analysis/output/tables/table_1_pharmacological_treatment.docx"))
}
```

```{r criminal history, message=FALSE, warning=FALSE, include=FALSE}
sentence_summary <-
  merged_data %>%
  filter(group == "patient") %>%
  select(age_first_crime_sentence, age_first_crime_reported, total_num_sentences) %>%
  psych::describe(ranges = TRUE, check = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  select(rowname, n, mean, sd, median, min, max) %>%
  mutate_if(is.numeric, round, 2)

crime_summary <-
  merged_data %>%
  filter(group == "patient") %>%
  mutate(crimes_under_treatment = factor(crimes_under_treatment),
         deadly_violence = factor(deadly_violence),
         assault_inc_agg = factor(assault_inc_agg),
         sexual_crimes = factor(sexual_crimes)) %>%
  select(crimes_under_treatment, deadly_violence, assault_inc_agg, sexual_crimes) %>%
  melt()

deadly_violence <- 
  crime_summary %>% 
  group_by(deadly_violence) %>% 
  tally() %>%
  mutate(percent = n / 27 * 100) %>%
  mutate_if(is.numeric, round, 1) %>%
  as.data.frame()

assault_inc_agg <- 
  crime_summary %>% 
  group_by(assault_inc_agg) %>% 
  tally() %>%
  mutate(percent = n / 27 * 100) %>%
  mutate_if(is.numeric, round, 1) %>%
  as.data.frame()

sexual_crimes <- 
  crime_summary %>% 
  group_by(sexual_crimes) %>% 
  tally() %>%
  mutate(percent = n / 27 * 100) %>%
  mutate_if(is.numeric, round, 1) %>%
  as.data.frame()
```

The criminal history of the MDOs was heterogeneous, with a mean total number of `r sentence_summary[3, 3]` sentences (SD = `r sentence_summary[3, 4]`), ranging from `r sentence_summary[3, 6]` to `r sentence_summary[3, 7]`. The mean age at first sentencing was `r sentence_summary[1, 3]` years (SD = `r sentence_summary[1, 4]`, range = `r sentence_summary[1, 6]` to `r sentence_summary[1, 7]`; note that 15 is the minimum age for criminal sentencing in Sweden), while the mean age at first reported crime was `r sentence_summary[2, 3]` years (SD = `r sentence_summary[2, 4]`, range = `r sentence_summary[2, 6]` to `r sentence_summary[2, 7]`). 

As per inclusion criteria, all MDOs had at some point been sentenced for a violent crime. `r deadly_violence[2, 2] + deadly_violence[3, 2]` MDOs (`r deadly_violence[2, 3] + deadly_violence[3, 3]`%) had committed acts of deadly violence, of which `r deadly_violence[3, 2]` (`r deadly_violence[3, 3]`%) repeatedly (i.e., on two or more occasions). A total of `r assault_inc_agg[2, 2] + assault_inc_agg[3, 2]` MDOs (`r assault_inc_agg[2, 3] + assault_inc_agg[3, 3]`%) had committed assault, including aggravated, of which `r assault_inc_agg[3, 2]` (`r assault_inc_agg[3, 3]`%) had done so repeatedly. Finally, `r sexual_crimes[2, 2] + sexual_crimes[3, 2]` MDOs (`r sexual_crimes[2, 3] + sexual_crimes[3, 3]`%) had committed sexual crimes, of which `r sexual_crimes[3, 2]` (`r sexual_crimes[3, 3]`%) repeatedly.

### Group differences

See Table 2 in output directory.

```{r group difference models, message=FALSE, warning=FALSE, include=FALSE}
source(here("analysis/scripts/r/05_robust_group_differences.R"))
```

### Correlations

See Table 3 in output directory.

```{r correlation models, message=FALSE, warning=FALSE, include=FALSE}
source(here("analysis/scripts/r/06_robust_correlation_general_dis.R"))
source(here("analysis/scripts/r/07_robust_correlation_nogo_acc.R"))
source(here("analysis/scripts/r/08_correlation_tables.R"))
```

```{r exploratory analyses, message=FALSE, warning=FALSE, include=FALSE}
if (!file.exists(here("analysis/data/tmp/correlation_mean_nogo_rt_formatted.rds")) || overwrite == TRUE) {
  brms_robust_correlation <- readRDS(here("analysis/scripts/brms/brms_robust_correlation.rds"))
  correlation_mean_nogo_rt <-
    rbind(
      get_correlation(merged_data, "median_nogo_rt", "nogo_accuracy"),
      get_correlation(merged_data, "median_nogo_rt", "p3_latency"))

  correlation_mean_nogo_rt_formatted <- correlation_mean_nogo_rt %>%
    mutate_if(is.numeric, round, 2) %>%
    mutate(rho = paste0(rho, " [", rho_lower_90, ", ", rho_upper_90, "]")) %>%
    select(variable_2, rho, p_direction)

  saveRDS(correlation_mean_nogo_rt_formatted, here("analysis/data/tmp/correlation_mean_nogo_rt_formatted.rds"))
}

if (!file.exists(here("analysis/data/tmp/group_differences_nogo_rt_formatted.rds")) || overwrite == TRUE) {
  brms_robust_group_diff <- readRDS(here("analysis/scripts/brms/brms_robust_group_diff.rds"))
  group_differences_nogo_rt <- get_group_difference(merged_data, "median_nogo_rt")

  group_differences_nogo_rt_full <- 
    group_differences_nogo_rt %>%
  mutate(
    mean_whole_group = mean,
    range_whole_group = range,
    mean_controls = mean.1,
    range_controls = range.1,
    mean_patients = mean.2,
    range_patients = range.2) %>%
  select(-c(mean, range, mean.1, range.1, mean.2, range.2))

  group_differences_nogo_rt_formatted <- group_differences_nogo_rt_full %>%
    mutate_if(is.numeric, round, 2) %>%
    mutate(diff = paste0(diff, " [", diff_lower_90, ", ", diff_upper_90, "]"),
           delta_est = paste0(delta_est, " [", delta_est_lower_90, ", ", delta_est_upper_90, "]"),
           p_direction = paste0(round(p_direction * 100, 2), "%")) %>%
    select(variable,
           mean_whole_group, range_whole_group,
           mean_controls, range_controls,
           mean_patients, range_patients,
           diff, delta_est, p_direction)

  saveRDS(group_differences_nogo_rt_full, here("analysis/data/tmp/group_differences_nogo_rt_full.rds"))
  saveRDS(group_differences_nogo_rt_formatted, here("analysis/data/tmp/group_differences_nogo_rt_formatted.rds"))
}

group_difference_nogo_rt <- readRDS(here("analysis/data/tmp/group_differences_nogo_rt_formatted.rds"))
correlation_mean_nogo_rt_formatted <- readRDS(here("analysis/data/tmp/correlation_mean_nogo_rt_formatted.rds"))
```

### Exploratory analyses

MDOs had longer median NoGo RTs than controls, responding on average `r group_difference_nogo_rt[, "diff"]` ms later than controls, indicating a relatively robust and large ($\hat{\delta}$ = `r group_difference_nogo_rt[, "delta_est"]`) group difference. Furthermore, median NoGo RT showed positive, moderate, and robust correlations with both NoGo accuracy (⍴ = `r correlation_mean_nogo_rt_formatted[1, 2]`, P~D~ = `r correlation_mean_nogo_rt_formatted[1, 3]`%) and NoGo P3 latency (⍴ = `r correlation_mean_nogo_rt_formatted[2, 2]`, P~D~ = `r correlation_mean_nogo_rt_formatted[2, 3]`%).

## Figures

See output directory.

```{r figures, message=FALSE, warning=FALSE, include=FALSE}
source(here("analysis/scripts/r/figure_1C.R"))
source(here("analysis/scripts/r/figure_2A.R"))
source(here("analysis/scripts/r/figure_3.R"))
source(here("analysis/scripts/r/figure_S1.R"))
source(here("analysis/scripts/r/figure_S2.R"))
```

# Session info

The statistical analysis was run on `r Sys.Date()`.

```{r echo=FALSE}
devtools::session_info()
```

